---
title: "STAT 4130 Project"
author: "Haoyang Ling"
date: "2023-07-22"
output:
  html_document:
    df_print: paged
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library("ggplot2")
library("GGally")
library(Hmisc)
library(leaps)
library(moments)
library(MASS)
library(car)
library(cowplot)
```

```{r}
### Load Data ###
diamonds <- read.csv("diamonds.csv")
diamonds <- subset(diamonds, select = -c(X))
diamonds$x_y = diamonds$x + diamonds$y
is_valid <- (diamonds$x > 0) & (diamonds$y > 0) & (diamonds$z > 0) & (diamonds$carat > 0)
diamonds <- diamonds[is_valid, ]
subset_diamonds = subset(diamonds, select = -c(cut, color, clarity))
```

```{r}
set.seed(413)
sample <- sample(c(TRUE, FALSE), nrow(diamonds), replace=TRUE, prob=c(0.7,0.3))
train  <- diamonds[sample, ]
test   <- diamonds[!sample, ]
write.csv(train, "train.csv")
write.csv(test, "test.csv")
```

```{r}
lm_full <- lm(price~., data = train)
summary(lm_full)
```

```{r}
y_train <- predict(lm_full, train)
mse <- mean((y_train - train$price)^2)
mse
```

```{r}
y_test <- predict(lm_full, test)
mse <- mean((y_test - test$price)^2)
mse
```

```{r}
ggplot(train, aes(x = price, y = y_train)) + 
geom_point() + 
geom_smooth(method = 'smooth') +
labs(x = "Actual Price", y = "Fitted Price")
```

### See the distribution of the dataset

```{r}
head(diamonds)
```

```{r}
hist(diamonds$price)
```
#### Comment:
In this picture, we can see that the price doesn't follow the normal distribution and therefore we may apply transformation to the price. A good choice is log, widely used to price.


#### Check the distribution of the variables

```{r}
hist.data.frame(subset(diamonds, select = -c(cut, color, clarity)))
```
#### Check Skewness

It seems that 

```{r}
skewness(subset_diamonds)
```

```{r}
diamonds$log_carat <- log(diamonds$carat)
diamonds$log_x <- log(diamonds$x)
diamonds$log_y <- log(diamonds$y)
diamonds$log_z <- log(diamonds$z)
diamonds$log_x_y <- log(diamonds$x_y)
skewness(subset(diamonds, select = -c(cut, color, clarity)))
```

#### Use the boxcox to transform variables into normal ones
```{r}
par(mfrow=c(3,3))
boxcox(lm(carat ~ 1, data=diamonds), xlab='carat lambda')
boxcox(lm(depth ~ 1, data=diamonds), xlab='depth lambda')
boxcox(lm(table ~ 1, data=diamonds), xlab='table lambda')
boxcox(lm(x ~ 1, data=diamonds), xlab='x lambda')
boxcox(lm(y ~ 1, data=diamonds), xlab='y lambda')
boxcox(lm(z ~ 1, data=diamonds), xlab='z lambda')
boxcox(lm(x_y ~ 1, data=diamonds), xlab='x_y lambda')
boxcox(lm(price ~ log_carat + log_x + log_y + log_z + log_x_y, data=diamonds))
diamonds$log_price <- log(diamonds$price)
```

```{r}
##ggpairs(subset(diamonds, select = -c(cut, color, clarity, carat, price, x, y, z, x_y)))
```
#### Comment
There is multi-colinearity in the dataset.

```{r}
vif(lm(log_price ~ log_carat + log_x + log_y + log_z, data=diamonds))
```
```{r}
res_c = lm(log_price ~ log_carat, data = diamonds)$residuals
res_x = lm(log_x ~ log_carat, data = diamonds)$residuals
res_y = lm(log_y ~ log_carat, data = diamonds)$residuals
res_z = lm(log_z ~ log_carat, data = diamonds)$residuals
res_x_y = lm(log_x_y ~ log_carat, data = diamonds)$residuals
par(mfrow=c(1,4))
plot(y=res_c, x=res_x) + abline(v = 0, col="red")
plot(y=res_c, x=res_y) + abline(v = 0, col="red")
plot(y=res_c, x=res_z) + abline(v = 0, col="red")
plot(y=res_c, x=res_x_y) + abline(v = 0, col="red")
```
```{r}
ggplot(data=diamonds, aes(x = clarity, y = log_price)) + geom_boxplot()
ggplot(data=diamonds, aes(x = color, y = log_price)) + geom_boxplot()
ggplot(data=diamonds, aes(x = cut, y = log_price)) + geom_boxplot()
```
```{r}
ggplot(data=diamonds, aes(x=log_carat, y=log_price, color=cut)) + geom_point(alpha=0.5) + geom_smooth(method = 'lm', formula = 'y ~ x')

ggplot(data=diamonds, aes(x=log_carat, y=log_price, color=clarity)) + geom_point(alpha=0.5) + geom_smooth(method = 'lm', formula = 'y ~ x')


ggplot(data=diamonds, aes(x=log_carat, y=log_price, color=color)) + geom_point(alpha=0.5) + geom_smooth(method = 'lm', formula = 'y ~ x')
```
#### Comment
It shows that the cut and clarity may have the interaction with log_carat, while color not.


```{r}
diamonds$color = factor(diamonds$color)
diamonds$cut = factor(diamonds$cut)
diamonds$clarity = factor(diamonds$clarity)
diamonds$cut_fair = ifelse(diamonds$cut == "Fair", 1, 0)
diamonds$clarity_I1 = ifelse(diamonds$clarity == "I1", 1, 0)
```


```{r}
ggplot(diamonds, aes(x = log_carat, y = log_price, color=color)) + 
geom_point() + 
facet_grid(cut_fair ~ clarity) + 
geom_smooth(method = 'lm') +
labs(x = "Log Carat", y = "Log Price")
```


```{r}
set.seed(413)
sample <- sample(c(TRUE, FALSE), nrow(diamonds), replace=TRUE, prob=c(0.7,0.3))
train  <- diamonds[sample, ]
test   <- diamonds[!sample, ]
write.csv(train, "train.csv")
write.csv(test, "test.csv")
```

```{r}
lm_full <- lm(log_price~.-price, data = train)
summary(lm_full)
```
```{r}
y_train_full <- exp(predict(lm_full, train))
mse <- mean((y_train_full - train$price)^2)
mse
```

```{r}
y_test_full <- exp(predict(lm_full, test))
mse <- mean((y_test_full - test$price)^2)
mse
```

```{r}
ggplot(train, aes(x = price, y = y_train)) + 
geom_point() + 
geom_smooth(method = 'smooth') +
labs(x = "Actual Price", y = "Fitted Price")
```


```{r}
lm1 <- lm(log_price ~ log_carat, data = train)
summary(lm1)
```

```{r}
ggplot(data=diamonds, aes(x = clarity, y = log_price - 1.674*log_carat)) + geom_boxplot()
ggplot(data=diamonds, aes(x = color, y = log_price - 1.674*log_carat)) + geom_boxplot()
ggplot(data=diamonds, aes(x = cut, y = log_price - 1.674*log_carat)) + geom_boxplot()
```

```{r}
qqnorm(rstudent(lm1))
qqline(rstudent(lm1))
```
```{r}
crPlots(lm1)
```
```{r}
ggplot(data = train, aes(x = log_x, y = lm1$residuals)) +
geom_point() +
geom_smooth() +
geom_hline(yintercept = 0, color = 'red')
```

#### See the errors
```{r}
y_train1 <- exp(predict(lm1, train))
mse <- mean((y_train1 - train$price)^2)
mse
```


```{r}
y_test1 <- exp(predict(lm1, test))
mse <- mean((y_test1 - test$price)^2)
mse
```

#### Color
```{r}
lm2 <- lm(log_price ~ color + log_carat, data = train)
anova(lm2)
```
```{r}
qqnorm(rstudent(lm2))
qqline(rstudent(lm2))
crPlots(lm2)
```

```{r}
y_train2 <- exp(predict(lm2, train))
mse <- mean((y_train2 - train$price)^2)
mse
```
```{r}
y_test2 <- exp(predict(lm2, test))
mse <- mean((y_test2 - test$price)^2)
mse
```

```{r}
lm3 <- lm(log_price ~ color + log_carat + log_carat:cut_fair, data = train)
anova(lm3)
```
```{r}
qqnorm(rstudent(lm3))
qqline(rstudent(lm3))
```

```{r}
avPlots(lm3)
```

```{r}
y_train3 <- exp(predict(lm3, train))
mse <- mean((y_train3 - train$price)^2)
mse
```
```{r}
y_test3 <- exp(predict(lm3, test))
mse <- mean((y_test3 - test$price)^2)
mse
```



```{r}
lm4 <- lm(log_price ~ log_carat*clarity_I1*cut_fair + color, data = train)
summary(lm4)
```

```{r}
y_train4 <- exp(predict(lm4, train))
mse <- mean((y_train4 - train$price)^2)
mse
```
```{r}
y_test4 <- exp(predict(lm4, test))
mse <- mean((y_test4 - test$price)^2)
mse
```
```{r}
par(mfrow = c(2, 2))
plot(lm4)
```


```{r}
lm5 <- lm(log_price ~ log_carat*clarity*cut + color, data = train)
summary(lm5)
```

```{r}
y_train5 <- exp(predict(lm5, train))
mse <- mean((y_train5 - train$price)^2)
mse
```
```{r}
y_test5 <- exp(predict(lm5, test))
mse <- mean((y_test5 - test$price)^2)
mse
```
```{r}
par(mfrow = c(2, 2))
plot(lm5)
```

```{r}
lm6 <- lm(log_price ~ log_carat*clarity + log_carat*cut + color*clarity, data = train, offset = log_carat)
summary(lm6)
```
```{r}
y_train6 <- exp(predict(lm6, train))
mse <- mean((y_train6 - train$price)^2)
mse
```


```{r}
y_test6 <- exp(predict(lm6, test))
mse <- mean((y_test6 - test$price)^2)
mse
```



```{r}
lm7 <- lm(log_price ~ log_carat*clarity + log_carat*cut + color*clarity, data = train)
summary(lm7)
```

```{r}
y_train7 <- exp(predict(lm7, train))
mse <- mean((y_train7 - train$price)^2)
mse
```


```{r}
y_test7 <- exp(predict(lm7, test))
mse <- mean((y_test7 - test$price)^2)
mse
```

```{r}
par(mfrow = c(2, 2))
plot(lm7)
```
```{r}
confint(lm7)
```
```{r}
anova(lm7)
```
```{r}
ggplot(test, aes(x = price, y = y_test7)) + 
geom_point() + 
geom_smooth(method = 'smooth') +
labs(x = "Actual Price", y = "Fitted Price")
```


```{r}
dim(diamonds)
```
```{r}
vif(lm7)
```


```{r}
step(
     lm(log_price ~ log_carat, data=diamonds), 
     scope = log_price ~ log_carat + carat + log_carat*depth + table + log_carat*color + log_carat*clarity + log_carat*cut,
     direction="both", 
     test="F",
     k=log(53920)
)
```

```{r}
lm8 <- lm(log_price ~ log_carat*color + log_carat*clarity + log_carat*cut + color*clarity, data = train)
sum <- summary(lm8)$coefficients
write.csv(sum, "sum.csv")
```

```{r}
y_test8 <- exp(predict(lm8, test))
mse <- mean((y_test8 - test$price)^2)
mse
```
```{r}
train$table_2 = train$table^-2
test$table_2 = test$table^-2
```


```{r}
lm9 <- lm(log_price ~ log_carat + cut + color + clarity + depth + table_2 + log_x + log_y + log_z + log_x_y, data = train)
summary(lm9)
```

```{r}
y_test9 <- exp(predict(lm9, test))
mse <- mean((y_test9 - test$price)^2)
mse
```
```{r}
par(mfrow = c(2, 3))
plot(lm7, which=1)
plot(lm8, which=1)
plot(lm9, which=1)
plot(lm7, which=2)
plot(lm8, which=2)
plot(lm9, which=2)
```


```{r}
p1 <- ggplot(test, aes(x = price, y = y_test7)) + 
geom_point() + 
geom_smooth(method = 'lm', color='red') +
labs(x = "Actual Price", y = "Fitted Price")
p2 <- ggplot(test, aes(x = price, y = y_test8)) + 
geom_point() + 
geom_smooth(method = 'lm', color='red') +
labs(x = "Actual Price", y = "Fitted Price")
p3 <- ggplot(test, aes(x = price, y = y_test9)) + 
geom_point() + 
geom_smooth(method = 'lm', color='red') +
labs(x = "Actual Price", y = "Fitted Price")
plot_grid(p1, p2, p3, labels = c('M1', 'M2', 'M3'), label_size = 12, nrow=1)
```

```{r}
anova(lm8, lm7)
```
```{r}
par(mfrow = c(1, 3))
plot(lm7, which=5)
plot(lm8, which=5)
plot(lm9, which=5)
```

